Brief Announcement: An Improved Lower Bound for Dynamic Graph Partitioning

### 2020-01-27 paper/paper.tex
small changes for k=3

\begin{proof}
\mahmoud{A simpler proof but still not fully rigorous. E.g., what if ALG collocates on $S_1$? Easy technicalities most likely.}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
Consider any online algorithm ALG.
We construct the input sequence of requests for ALG as follows.
First,
we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
nodes on clusters $S_1$.
Note that this does not cost ALG.
Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
Next,
we issue the request between $x_0$ and $y_0$.
Since this is an external request,
ALG joins them into one component and collocates them in some cluster other than $S_1$.
For this,
ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
Next,
we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
consequently replaces $x_1$.
We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
 until there are only two single nodes left that  originate from the same cluster.
I.e.,
two single nodes $x^*, y^*,I(x^*) = I(y^*)$
that constitute the last pair of such nodes.
At this point there are at most $\ell+1$ single nodes left
since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
nodes.
Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
Hence,
OPT pays $4$ for node migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
\end{proof}

@@

\begin{theorem}
  There exists a $O(\ell)$-competitive algorithm for $k=3$.
\end{theorem}

\begin{proof}
  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist. Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
\end{proof}

We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
towards an algorithm for the general online problem.
Note that in the general model,
an optimal partition is not a necessarily a perfect partition.
We divide the sequence of requests into phases.
A phase terminates when there is no perfect partition of components.
We apply requests to $ON_{k=3}$ as before. 
Once the current phase terminates,
we reset all components by removing (forgetting) all revealed edges,
 and then we proceed to the next phase.

\begin{corollary} \label{cor:k=3}
    There exists $O(\ell)$-competitive  algorithm for the general online problem when $k=3$. 
\end{corollary}
\begin{proof}
    Consider any phase and assume it terminates upon arrival of some request $r$.
     By Lemma  \ref{lemma:k=3},
    ON pays $O(\ell)$ during the phase.
    Regardless of the of OPT's partition at the beginning of the phase,
    as long as OPT stays in the same configuration,
     it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
     Else,
     OPT incurs the cost of moving to new partitions.
     In any case,
     OPT must pay at least one during the phase.
\end{proof}

@@

\begin{appendix}

\end{appendix}

>>>>

\begin{proof}
We construct an instance of the problem with $\ell$ clusters 
$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
Consider any online algorithm ALG.
We construct the input sequence of requests for ALG as follows.
First,
we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
nodes on clusters $S_1$.
Note that this does not cost ALG.
Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
Next,
we issue the request between $x_0$ and $y_0$.
Since this is an external request,
ALG joins them into one component and collocates them in some cluster other than $S_1$.
For this,
ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
Next,
we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
consequently replaces $x_1$.
We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
 until there are only two single nodes left that  originate from the same cluster.
I.e.,
two single nodes $x^*, y^*,I(x^*) = I(y^*)$
that constitute the last pair of such nodes.
At this point there are at most $\ell+1$ single nodes left
since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
nodes.
Given this sequence of requests,
the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 swapping $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
Hence,
OPT pays $4$ for node migrations and
ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
\end{proof}

@@

   Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
on each inter-cluster request.
\begin{theorem}
	Algorithm \emph{Arb-partition} is $O(\ell)$-competitive when $k=3$.
\end{theorem}

\begin{proof}
	The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
	Each request triggers at most one re-balancing in \emph{Arb-partition}
	which costs $O(1)$ by Theorem~\ref{rebalancing-cost}.
	Then,
	the total cost of any algorithm is $O(\ell)$.
\end{proof}

We leverage the simple algorithm \emph{Arb-partition} for $k=3$
towards an algorithm for the general problem
where the communication graph is not restricted.
Note that in the general model,
an optimal partition is not a necessarily a perfect partition.
We divide the sequence of requests into phases.
A phase terminates when there is no perfect partition of components.
We apply requests to \emph{Arb-partition} as before. 
Once the current phase terminates,
we reset all components by removing (forgetting) all revealed edges,
and then we proceed to the next phase.

\begin{theorem} \label{cor:k=3}
	There exists $O(\ell)$-competitive algorithm for the general online problem when $k=3$. 
\end{theorem}
\begin{proof}
	Consider any phase and assume it terminates upon arrival of some request $r$.
	By Lemma  \ref{lemma:k=3},
	\emph{Arb-partition} pays $O(\ell)$ during the phase.
	If OPT stays in the same configuration throughout the phase,
	regardless of its partition at the beginning of the phase,        
	it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
	Else,
	OPT incurs the cost of moving to new partitions.
	In any case,
	OPT pays at least one during the phase and 
	the competitive ratio is $O(\ell)$.
\end{proof}

@@

\begin{appendix}
\end{appendix}
